---
title: "Visualizations - sales data"
output: md_document
always_allow_html: yes
#output: html_document 
---

```{r readLibraries, echo = FALSE, message=FALSE, warning=FALSE }

require(dplyr)
require(ggplot2)
require(GGally)
require(scales)
require(alluvial)
require(ggalluvial)
require(arules)
require(arulesViz)
require(tidyverse)
require(ca)
require(mlr)
library(data.table)
require(knitr)
require(kableExtra)
source("functions.R")

```


```{r readData, echo = FALSE, message = FALSE, warning = FALSE }

fileDataName <- "BlackFriday.csv"
pathData <- file.path(getwd(), "data")
fileData <- file.path(pathData, fileDataName)

data <- read.csv(file = fileData, header = TRUE, sep = ",")
descriptionsFeatures <- describe_variables(data)

```

# Data structure

In dataset are 12 columns. The dataset contains different kinds of variables - numerical or categorical.
I used my function `DescribeVariables()` to get more infromations about structure of data. You can find code of this function in script prepare_data_func.R  
Here the results:
<br/>
<br/>

```{r features, echo = FALSE, message = FALSE, warning = FALSE }

kable(descriptionsFeatures, booktabs = T, caption = "Table 1: Structure of data") %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  column_spec(1, width = "8cm", color = "red", bold = TRUE) %>%
  row_spec(0, bold = TRUE)

```
<br/>
<br/>

Type of variables: *Occupation*, *Product_Category_1*, *Product_Category_2* and *Marital_Status* is integer. It needed to conver types this variables to factor.

```{r converTypeVariables, echo = TRUE, message = FALSE, warning = FALSE }

data$Occupation         <- factor(data$Occupation,         levels = sort(unique(data$Occupation)))
data$Product_Category_1 <- factor(data$Product_Category_1, levels = sort(unique(data$Product_Category_1)))
data$Product_Category_2 <- factor(data$Product_Category_2, levels = sort(unique(data$Product_Category_2)))
data$Product_Category_3 <- factor(data$Product_Category_3, levels = sort(unique(data$Product_Category_3)))
data$Marital_Status     <- factor(data$Marital_Status,     levels = sort(unique(data$Marital_Status)))


```

```{r saveRdata, echo = FALSE, message = FALSE, warning = FALSE }
if(!file.exists(file.path(pathData, "data.Rdata")))
  save(data, file = file.path(pathData, "data.Rdata"))

```

<br/>
Checking for NA values in columns we can see that the column *Product_Category_2* is fill by 31% of NA values and the column *Product_Category_3* is fill by 69% NA values. We have got colum User_ID to identify each customer. We have got 3 623 unique products in data (*Product_ID*). 

```{r dataTransformation, echo = FALSE, message = FALSE, warning = FALSE }

theMostChosenProductCustomer <- data %>%
  select( c("User_ID", "Product_Category_1", "Purchase")) %>%
  group_by(User_ID, Product_Category_1) %>%
  add_count(Product_Category_1) %>%
  group_by(User_ID) %>%
  slice(which.max(n))%>%
  plyr::rename(c("n" = "numberOfProduct1", "Product_Category_1" = "mostChosenProduct1", "Purchase" = "allPurchaseMostChosenProduct1")) %>%
  arrange(desc(numberOfProduct1))


dataCustomer <- data %>%
  select(-c("Product_ID","Product_Category_1","Product_Category_2","Product_Category_3")) %>%
  dplyr::group_by(User_ID, Gender, Age, Occupation, City_Category, Stay_In_Current_City_Years, Marital_Status) %>%
  dplyr::summarize(sumOfPurchase = sum(Purchase, na.rm = TRUE),
                   meanOfPurchase = mean(Purchase, na.rm = TRUE),
                   medianOfPurchase = median(Purchase),
                   numberOfTransactions = n(),
                   minPurchase = min(Purchase),
                   maxPurchase = max(Purchase)) %>%
  left_join(theMostChosenProductCustomer, by = "User_ID") 
  
dataCustomer$mostChosenProduct1 <- factor(dataCustomer$mostChosenProduct1, levels = c(1,2,3,4,5,6,7,8,10,11,12,13,15,16,18))


save(dataCustomer, file = file.path(pathData, "dataCustomer.Rdata"))
```
<br/>
<br/>
<br/>
<br/>

# Exploratory Data Analysis
<br/>
<br/>

## Purchase
<br/>
<br/>
```{r visualizationsData, echo = FALSE, message = FALSE, warning = FALSE}

ggplot(data, aes(Purchase))+
  geom_histogram(aes(y=..density..),colour="black", fill="#FF6666",binwidth = 1000)+
  ggtitle("Histogram of purchase")
  

ggplot(data, aes(City_Category, Purchase, fill=City_Category))+
  geom_boxplot( outlier.colour = "red", outlier.shape = 1)+
  coord_flip()+
  labs(title = "Boxplots of Purchase by City Category")+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))

data %>%
  select(c("Marital_Status", "Stay_In_Current_City_Years", "Purchase")) %>%
  group_by(Marital_Status, Stay_In_Current_City_Years) %>%
  summarize(Purchase = dollar(round(sum(Purchase)/1000000, 0), suffix = "mln")) %>%
  ggplot(aes(x = Marital_Status, y = Purchase, fill = Marital_Status)) +
  geom_text(aes(label = Purchase), position = position_dodge(width = 0.7), vjust = -1, hjust = 0.5, size = 3)+
  geom_bar( stat = "identity", position = "dodge2", width = 0.4) +
  labs(y = "Purchase mln", subtitle = "Stay_In_Current_City_Years") +
  theme(plot.subtitle = element_text(hjust = 0.5)) +
  facet_grid( ~ Stay_In_Current_City_Years)


ggplot(dataCustomer, aes(x = Gender,y = sumOfPurchase/1000, fill = Gender ))+
  geom_boxplot()+
  labs(y = " Customers's purchase in thousands $")+
  facet_grid( ~ Age)+
  labs(title = "Age vs Gender")+
  theme(plot.title = element_text(hjust = 0.5))


data %>%
  select(c("Age", "City_Category", "Purchase", "Gender")) %>%
  group_by(Age, City_Category, Gender)%>%
  summarize(Purchase = (round(sum(Purchase)/1000000, 0))) %>%
  ggplot(aes(y = Purchase, axis1 = Age, axis2 = City_Category)) +
  geom_alluvium(aes(fill = City_Category), width = 1/13) +
  geom_stratum(width = 1/13, fill = "black", color = "grey") +
  geom_label(stat = "stratum", label.strata = TRUE) +
  scale_x_discrete(limits = c("Age", "City_Category"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle("Purchase  by Age and City Category") +
    theme(legend.position = "none", plot.title = element_text(hjust = 0.5))


data %>% 
  select(c("Stay_In_Current_City_Years", "City_Category", "Purchase", "Gender")) %>%
  group_by(Stay_In_Current_City_Years, City_Category,Gender) %>%
  summarize(Purchase = (round(sum(Purchase)/1000000, 0))) %>%
  ggplot(aes(y = Purchase, axis1 = Stay_In_Current_City_Years, axis2 = City_Category)) +
  geom_alluvium(aes(fill = City_Category), width = 1/13)+
  geom_stratum(width = 1/13, fill = "navy", color = "grey")+
  geom_label(stat = "stratum", label.strata = TRUE) +
  scale_x_discrete(limits = c("Stay_In_Current_City_Years", "City_Category"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle("Purchase by Stay_In_Current_City_Years and City Category")+
  theme(plot.title = element_text(hjust = 0.5))

```
<br/>
<br/>
<br/>
<br/>


## Customers
<br/>
<br/>

```{r plotsCustomers, echo = FALSE, message = FALSE, warning = FALSE}
ggplot( data = dataCustomer, aes(x = Age, fill = Gender))+
  geom_bar(data = subset(dataCustomer, Gender == "F"))+
  geom_bar(data = subset(dataCustomer ,Gender == "M"), aes(y=..count..*(-1)))+
  scale_y_continuous(breaks = seq(-1500,1500,500),labels=abs(seq(-1500,1500,500))) + 
  coord_flip() +
  scale_fill_brewer(palette = "Dark2") +
  ggtitle("Pyramid plot of Age and Gender Customers") +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(dataCustomer, aes(x = Gender, y=count(), fill = Marital_Status))+
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = percent)+
  ylab("")

ggplot(dataCustomer, aes(numberOfTransactions)) +
  geom_density(fill = 'cyan') +
  ggtitle("Density of customers transactions") +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(dataCustomer, aes(numberOfTransactions))+
  geom_density(fill = 'cyan')+
  facet_wrap( ~ Age) +
  ggtitle("Density of customers transactions in each group of age")+
  theme(plot.title = element_text(hjust = 0.5))
  

dataCustomer[ ,c("Age", "numberOfTransactions")] %>%
  group_by(Age)%>%
  summarise(sumOfTransactions = sum(numberOfTransactions))%>%
  ggplot( aes(x = Age, y = sumOfTransactions, fill = Age)) +
  geom_bar(stat="identity", position = "stack") +
  labs(title = "Age vs sum of transactions") +
  theme(plot.title = element_text(hjust = 0.5))


dataCustomer %>%
  select(c("Age", "numberOfTransactions", "Marital_Status", "Gender")) %>%
  mutate(Marital_Status = paste0("marital status:", Marital_Status)) %>%
  ggplot( aes(Age, numberOfTransactions, fill=Age))+
  stat_boxplot(geom="boxplot", position = "dodge" )+
  facet_grid(Gender ~ Marital_Status)+
  theme(strip.background = element_rect(colour = "black", fill = "white"))  


ggplot(dataCustomer, aes(numberOfTransactions))+
  geom_density(aes(color = Gender))+
  theme(legend.position = "none")+
  facet_wrap( ~ Gender)

```
<br/>
<br/>
<br/>
<br/>


## Products
<br/>
<br/>
<br/>

```{r products, echo = FALSE, message = FALSE, warning = FALSE }

kable(data %>%
        select(c("Product_ID", "Product_Category_1")) %>%
        dplyr::group_by(Product_Category_1)%>%
        dplyr::summarize(numberOfProducts = n())%>%
        mutate(percentage = round(numberOfProducts/sum(numberOfProducts),2))%>%
        arrange(desc(numberOfProducts)) %>% head(10), format="markdown")  %>%
  kable_styling( "striped", full_width = FALSE, position = "center") %>%
  column_spec(c(1, 2, 3), width = "4cm") %>%
  row_spec(0, bold = TRUE)

kable(data %>%
        select(c("Product_Category_2","Product_ID")) %>%
        dplyr:: group_by(Product_Category_2) %>%
        dplyr:: summarise(numberOfProducts = n()) %>%
        mutate(percentage = round(numberOfProducts / sum(numberOfProducts),2)) %>%
        arrange(desc(numberOfProducts)) %>% head(10), format="markdown")%>% 
  kable_styling( "striped", full_width = FALSE, position = "center") %>%
  column_spec(c(1, 2, 3), width = "4cm") %>%
  row_spec(0, bold = TRUE)

kable(data %>%
        select(c("Product_Category_3","Product_ID")) %>%
        dplyr:: group_by(Product_Category_3) %>%
        dplyr:: summarise(numberOfProducts = n()) %>%
        mutate(percentage = round(numberOfProducts / sum(numberOfProducts),2)) %>%
        arrange(desc(numberOfProducts)) %>% head(10), format="markdown") %>% 
  kable_styling( "striped", full_width = FALSE, position = "center") %>%
  column_spec(c(1, 2, 3), width = "4cm") %>%
  row_spec(0, bold = TRUE)




colors<-grDevices::colors()

 with(data,
      spineplot(x = Product_Category_1, y = Age, main = "Prefered Product Category 1 by Age Cutomer", col = sample(colors, 5)))


 with(data,
      spineplot(x = Product_Category_1, y = City_Category,
                main = "Prefered Product Category 1 by City Category", col = sample(colors, 5)))


 with(data,
      spineplot(x = Product_Category_1, y = Gender,
                main = "Prefered Product Category 1 by Gender", col = sample(colors, 5)))

 with(data,
      spineplot(x = Product_Category_1, y = Marital_Status,
                main = "Prefered Product Category 1 by Marital Status", col = sample(colors, 5)))


 with(data,
      spineplot(x = Product_Category_1, y = Stay_In_Current_City_Years,
                main = "Prefered Product Category 1 by Stay_In_Current_City_Years", col = sample(colors, 5)))
```
