---
title: "Analysis- sales data"
output: md_document 
---

```{r readLibraries, echo = FALSE, message=FALSE, warning=FALSE }
require(dplyr)
require(ggplot2)
require(arules)
require(arulesViz)
require(tidyverse)
require(ca)
load(file.path(getwd(), "data","data.Rdata"))
```
<br/>
<br/>

## <span style="color:blue"> *Market Basket Analysis* </span>

### **Apriori Algorithm** 
Using this Apriori algorithm we can predict what products customers buy together. 
This is Associations Rule Mining.
The purpose of using Associations Rule Mining is found in retailing, clustering and classification.

Store owners can use association rules for numerous marketing strategies for example:  
* changing the store layout,  
* customer behavior analysis,  
* what are the trending items customers buy?  

### Application in R 
Firstly we have to get information about transactions from data. Each transaction shows items bought in that transaction and wchich User ID. 

#### **Prepare data**
First it's important to convert data frame into transaction data (transaction object --> `basket`). 
```{r prepareDataAprori, echo = TRUE, message = FALSE, warning = FALSE }
products     <- data$Product_ID
users        <- as.factor(data$User_ID)
sparse_matix <- split(products, users)
basket       <- as(sparse_matix, "transactions")
```

Here some information about new object `basket`. There are 5891 transactions (rows) and 3623 items (columns).
```{r objectAprori, echo = TRUE, message = FALSE, warning = FALSE }
summary(basket)
```
<br/>
Below the *Absolute Item Frequency Plot* presents top 10 highest frequency products bought.
```{r analysisOfAproriPlots, echo = FALSE, message = FALSE, warning = FALSE }
colors<-grDevices::colors()
itemFrequencyPlot(basket, topN = 10, type = "absolute", col = colors[10:19], main = "Absolute Item Frequency Plot")
```
<br/>

#### **Generating rules**
Create rules by function `apriori`. 
See the strongest association rules. There are 6079 rules. The rules inform what customers purchase if they buy another product.

*Support* and *Confidence* measure how interesting the rule is.  
*Lift* gives information about correlation between for example Produkct 1 and Product2 in the rule Product1=> Product2.  
This measure shows how one item (Product1) effects the onther item (Product2).
```{r generatingRules, echo = TRUE, message = FALSE, warning = FALSE }
rules <- apriori(data = basket, parameter = list(support = 0.004, confidence = 0.75) )
```

#### **Results**

Look at the top 10 sorted rules by decreasing metric *LIFT*.   
The first rule is: If customers buy product P00156442 and P00219042 they will also buy product P00304142 in 80 % in the cases.  
The second rule is: If customers buy product P00244242 and P00259042 they also buy product P00244142 in 77% in the cases, etc.  
```{r results, echo = FALSE, message = FALSE, warning = FALSE}
inspect(sort(rules, by = "lift")[1:10, ])
```
<br/>

#### **Visualizing Association Rules**

The Scatter plot rules shows that rules with hight *lift* have low *support*.
```{r visual, echo = FALSE, message = FALSE, warning = FALSE}
plot(rules)

#inspectDT(rules)
```
<br/>

```{r graph, echo = FALSE, message = FALSE, warning = FALSE}
topRules <- head(rules, n=10, by= "confidence")
plot(topRules, method = "graph")

plot(topRules, method="paracoord")
```



## <span style="color:blue"> *Correspondence Analysis* </span>

### **Correspondece analysis (CA)** 
Correspondence analysis is a popular tool for visualizing. This analysis applies to categorical rather than continuous data. This technique s an appropriate technique to explore relationships amongst variable response categories.

It's nedeed to prepare [(contingenty tables)](https://en.wikipedia.org/wiki/Correspondence_analysis).


```{r, tables, echo = TRUE, message = FALSE, warning = FALSE }

load(file.path(getwd(), "data","dataCustomer.Rdata"))

dataCustomer <- dataCustomer %>%
  select(c("Gender", "Age", "Occupation", "City_Category", "Stay_In_Current_City_Years", "Marital_Status", "sumOfPurchase"))%>%
  mutate(purchaseQuantile = cut(sumOfPurchase, breaks = quantile(dataCustomer$sumOfPurchase, probs = seq(0, 1, 0.25))),
         purchaseQuantile = purchaseQuantile/1000)

dataCustomer <- dataCustomer %>%
  mutate(purchaseQuantile = cut(sumOfPurchase, breaks = quantile(dataCustomer$sumOfPurchase, probs = seq(0, 1, 0.25)), include.lowest = TRUE))

dataCustomer$purchaseQuantile <-  case_when(dataCustomer$purchaseQuantile == "(566,6.92e+03]" ~ "566$-6920$",
                                        dataCustomer$purchaseQuantile == "(6.92e+03,8.6e+03]" ~ "6920$-8600$",
                                        dataCustomer$purchaseQuantile == "(8.6e+03,1.21e+04]" ~ "8600$-12100$",
                                        dataCustomer$purchaseQuantile ==  "(1.21e+04,2.39e+04]" ~ "12100$-23900$",
                                        TRUE ~ as.character(dataCustomer$purchaseQuantile))

dataCustomer$purchaseQuantile <- factor(dataCustomer$purchaseQuantile, levels = sort(unique(dataCustomer$purchaseQuantile)))


tab1 <- with(dataCustomer, table(Age, purchaseQuantile))
chisq.test(tab1)
plot(ca(tab1), arrows = c(TRUE, FALSE))
summary(ca(tab1))


tab2 <- with(dataCustomer, table(City_Category, purchaseQuantile ))
chisq.test(tab2)
plot(ca(tab2), arrows = c(TRUE, FALSE))
summary(ca(tab2))


```







